---
title: "3D Models from mosaics & rayshader"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{models-from-other}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
rgl::setupKnitr()
```

```{r setup, include = FALSE}
library(brickr)
```

## Getting started

The `bricks_from_*` series of functions creates 3D models of LEGO bricks from a variety of input formats.

## 3D mosaics

Begin with a brickr mosaic from an image. Rather than graphically rendering the mosaic using `build_mosaic()`, use `bricks_from_mosaic()`. This function takes two other inputs:

* `mosaic_height` is the number of bricks stacked at the mosaic's highest point. The default is 6.
* `highest_el` specifies if 'light' or 'dark' color bricks should be the tallest in the model. The default is 'light'. 

```{r bricks_6, rgl=TRUE, dev='png', echo=TRUE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4}
demo_img = tempfile() 
download.file("http://ryantimpe.com/files/mf_unicorn.PNG", demo_img, mode="wb")

mosaic <- png::readPNG(demo_img) %>% 
  image_to_mosaic()

mosaic %>% build_mosaic()
```

```{r bricks_6a, rgl=TRUE, dev='png', echo=TRUE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4}
mosaic %>% 
  bricks_from_mosaic(highest_el = "dark") %>% 
  build_bricks()

#From dput(round(rgl::par3d("userMatrix"),1)) after manual rotation
custom_rotation <- structure(c(0.9, 0.3, -0.3, 0, -0.3, 0.9, -0.3, 
                               0, 0.2, 0.4, 0.9, 0, 0, 0, 0, 1), .Dim = c(4L, 4L))

rgl::par3d(userMatrix = rgl::rotate3d(custom_rotation, 0, 0, pi/4 ,1))
```

## Models from rayshader

[rayshader](https://www.rayshader.com/) by [Tyler Morgan-Wall](https://twitter.com/tylermorganwall) is an open source package for producing 2D and 3D data visualizations in R. rayshader uses elevation data in a base R matrix and a combination of raytracing, spherical texture mapping, overlays, and ambient occlusion to generate beautiful topographic 2D and 3D maps. (Note: text lifted straight from [rayshader.com](https://www.rayshader.com/).)

3D models in **brickr** are rendered using the functions in **rayshader**. Using `bricks_from_rayshader()`, you can convert rayshader map output into a brickr model. This function takes three inputs:

* `hillshade` is topographic image matrix with an RGB channel (much like the mosaics).
* `heightmap` is a two-dimensional matrix specifiying the height of the image at each location.
* `max_height` is the number of bricks stacked at the mosaic's highest point. The default is 12.
* `img_size` is the number of bricks on each side of the model. The default is 48.

```{r bricks_rayshader, echo=TRUE, warning=FALSE, message=FALSE, fig.width=4, fig.height=4}
library(rayshader)

#Example from rayshader.com

#Here, I load a map with the raster package.
loadzip = tempfile() 
download.file("https://tylermw.com/data/dem_01.tif.zip", loadzip)
localtif = raster::raster(unzip(loadzip, "dem_01.tif"))
unlink(loadzip)

#And convert it to a matrix:
elmat = matrix(raster::extract(localtif, raster::extent(localtif), buffer = 1000),
               nrow = ncol(localtif), ncol = nrow(localtif))

rayshader_object <- elmat %>%
  sphere_shade(texture = "desert") %>%
  add_water(detect_water(elmat), color = "desert") %>%
  add_shadow(ray_shade(elmat, zscale = 3, maxsearch = 300), 0.5) 

#Plot with rayshader
rayshader_object %>%
  plot_3d(elmat, zscale = 10, fov = 0, theta = 135, zoom = 0.75, phi = 45, windowsize = c(1000, 800))

rayshader::render_snapshot(clear = TRUE)

#Plot as bricks
rayshader_object %>% 
  bricks_from_rayshader(elmat) %>% 
  build_bricks_rayshader(theta = 135, phi = 45)

rayshader::render_snapshot(clear = TRUE)
```

This example is rendered using `build_bricks_rayshader()`, which a bit faster for very large sets.